Fusion Moves for Correlation Clustering

### 2015-04-06 paper/cvpr15_abstract_mine.tex
added extended abstract

%!TEX root = ./egpaper_for_review.tex
%
\section{Introduction}
Correlation clustering~\cite{Bansal-2002}, also known as the multicut problem~\cite{chopra_1993_mp} 
is a basic primitive in computer vision~\cite{andres_2011_iccv,kroeger_2012_eccv,yarkony_2012_eccv,alush_2013_simbad} and data mining~\cite{Arasu-2009,Sadikov-2010,Chen-2012,Chierichetti-2014}.
See Sec.~\ref{sec:problem_formulation} for its formal definition of clustering the nodes of a graph.

@@

% Importantly, this allows doing justice to evidence in the data that two nodes or pixels do \emph{not} wish to end up in the same cluster or segment. This is in contrast to the submodular potentials so popularized by the graph cut algorithm, which only allow two nodes to be attracted to each other, or at most be agnostic about membership in the same cluster. Secondly, the algorithm does not require a specification of the number of clusters. This is in contrast to methods such as normalized cut \cite{} that can only accommodate attractive interactions and hence need the number of clusters to be specified.

% \paragraph{Contribution.} The evident usefulness of correlation clustering and its clean and compact formulation in terms of an optimization problem, together with its unfortunate NP-hardness, are an invitation to develop fast approximate solvers. The basic idea of the move making algorithm proposed here is to maintain, at all times, a best current partitioning; and to iteratively improve it (as we show, monotonously) by considering diverse and cheaply generated proposal partitionings. Any two nodes that are in one cluster in both the current and the proposal partitioning are contracted. Edges to contracted nodes are equally combined and reweighted, and the correlation clustering problem is then solved on this reduced graph. We offer a polyhedral characterization of this strategy, and evaluate it in conjunction with two versatile proposal generators.   We conduct experiments on a broad range of data sets ranging from 2D and 3D segmentation problems to clustering in signed social networks. The results suggest that this simple move making algorithm is the fastest technique known today, reaching close to globally optimal solutions in a tenth or a hundredth of the time required by the most efficient exact solvers. 
\vspace{0.1cm}
\noindent \textbf{Contribution.}
In this work we present novel approaches that are designed for large scale correlation clustering problems.
First, we define a novel energy based agglomerative clustering algorithm that monotonically increases the energy.
With this at hand we show how to improve the anytime performance of Cut, Clue \& Cut~\cite{beier_2014_cvpr}.
%
Second, we improve the anytime performance of polyhedral multicut methods~\cite{kappes_2013_arxiv} by more efficient separation procedures.
%
Third, we introduce cluster-fusion moves, which extend the original fusion moves~\cite{Lempitsky-2010} 
used in supervised segmentation to the unsupervised case and give a polyhedral interpretation of this algorithm.
Finally, we propose two versatile proposal generators, and evaluate the proposed methods on existing and new benchmark problems.
Experiments show that we can improve the computation time by one to two magnitudes without worsening the segmentation 
quality significantly.

\vspace{0.1cm}
\noindent \textbf{Related Work.}
A natural approach is to solve the integer linear program (ILP) directly~\ref{eq:edgeproblem}. 
To this end, efficient separation procedures have been found~\cite{kappes_2011_emmcvpr,kappes_2013_arxiv} that allow to iteratively augment the set of constraints until a valid partitioning is found. 
Alternatively, it is possible to relax the integrality constraints of the ILP formulation~\cite{kappes_2013_arxiv}. 
Such an outer relaxation can be iteratively tightened. However, intermediate solutions are fractional and therefore rounding is required to obtain a valid partitioning.
For the latter approach column generating methods exist, which work best on planar graphs~\cite{yarkony_2012_eccv}. %or almost planar \cite{yarkony-andres-2013} graphs.

Another line of work uses move making algorithms 
to optimize correlation clustering~\cite{Kernighan-1970,bagon_2011_arxiv,beier_2014_cvpr}.
Starting with an initial segmentation, auxiliary max-cut problems are (approximately) solved,
such that the segmentation is strictly improved.
As shown in~\cite{beier_2014_cvpr} only Cut, Glue \& Cut (CGC)
can deal with large scale problems, but can also suffer from very large auxiliary problems.

Outside computer vision, greedy methods~\cite{Soon-2001,Ng-2002,Gionis-2007,Elsner-2008,Ailon-2008} have been suggested for correlation clustering problems, see \cite{Elsner-2009} for an overview.
The PIVOT Algorithm~\cite{Ailon-2008} iterates over all nodes in random order.
If the node is not assigned it constructs a cluster containing the node and all its 
unassigned positively linked neighbors.  
%
A widely used post-processing method is  Best One Element Move (BOEM)~\cite{Gionis-2007}, which iteratively reassigns nodes to clusters.

For energy minimization problems fusion moves have become increasingly popular~\cite{Lempitsky-2010,kappes_2014_ws}.
For many large scale computer vision applications fusion moves lead to good approximations
with state of the art anytime performance~\cite{kappes_2014_ws}.
Due to the ambiguity of a node-labeling, classical fusion moves~\cite{Lempitsky-2010} cannot be applied directly for correlation clustering.
We will show how to overcome this problem in Sec.~\ref{sec:cc_fm}.

\vspace{0.1cm}
\noindent \textbf{Outline:} 
In Sec.~\ref{sec:problem_formulation} we give a 
detailed problem definition and introduce 
the correlation clustering objective.
Next we give a description of energy based hierarchical clustering in Sec.~\ref{sec:ehc} and
our proposed correlation clustering fusion moves in Sec.~\ref{sec:cc_fm}.
We evaluate the proposed methods in Sec.~\ref{sec:exp} and conclude in Sec.~\ref{sec:future} and \ref{sec:conclusion}.

%-------------------------------------------------------------------------

\input{inputs/fig_notation.tex}
\section{Notation and Problem Formulation}\label{sec:problem_formulation}
Let $G=(V,E, w)$ be a weighted graph of nodes $V$ and edges $E$.
%
The function $w : E \rightarrow \mathbb{R}$ assigns a weight to each edge.
We will use $w_e$ as a shorthand for $w(e)$.
A positive weight expresses the desire that two adjacent nodes should
be merged, whereas a negative weight indicates
that these nodes should be separated into two distinct regions.
%
%A \emph{subgraph} $G_A = \{A, E_A, w\}$ consists
%of nodes $A \subseteq V$ and edges $E_A := E\cap (A\times A)$.
%
%\input{inputs/fig_notation.tex}
A segmentation of the graph $G$ can be either given by a
node labeling $l \in \mathbb{N}^{|V|}$
or an edge labeling $y \in\{0,1\}^{|E|}$, \cf Fig.~\ref{fig:notation}.
An edge labeling is only consistent if it does not violate any cycle constraint~\cite{chopra_1993_mp}.
We denote the set of all consistent edge labelings by $P(G)\subset\{0,1\}^{|E|}$.
The convex hull of this set is known as the \emph{multicut polytope} $MC(G) = \textrm{conv}(P(G))$.
%
By $l(y)$ we denote some node labeling for a segmentation given by $y$.

Given a weighted graph $G=(V,E,w)$ we consider the problem of segmenting $G$ such that the costs
of the edges between distinct segments is minimized. This can be formulated in the node domain
by assigning each node $i$ a label $l_i \in \mathbb{N}$
\begin{align}
  l^* &= \argmin_{l \in \mathbb{N}^{|V|}} \sum_{ (i,j) \in E } w_{ij} \cdot [l_{i} \neq l_{j}], \label{eq:nodeproblem}
\end{align} 
or in the edge domain, by labeling each edge $e$ as cut $y_e=1$ or uncut $y_e=0$ 
\begin{align}
  y^* &= \argmin_{y \in P(G)} \sum_{ (i,j) \in E } w_{ij} \cdot y_{ij} \label{eq:edgeproblem}.%\\ 
\end{align}
As shown in~\cite{kappes_2013_arxiv}  both problems are equivalent, but formulation \ref{eq:nodeproblem}
suffers from ambiguities in the representation, \cf Fig.~\ref{fig:notation}.

%\input{inputs/fig_notation.tex}

>>>>

% Use class option [extendedabs] to prepare the 1-page extended abstract.
\documentclass[extendedabs]{bmvc2k}
\usepackage[colorlinks = true,
            linkcolor = blue,
            urlcolor  = blue,
            citecolor = blue,
            anchorcolor = blue]{hyperref}

% Document starts here
\begin{document}

\title{Fusion Moves for Correlation Clustering}
\addauthor{
Thorsten Beier$^1$, 
Fred A. Hamprecht$^2$
J\"org H. Kappes$^3$
}{}{1}
\addinstitution{
$^1$thorsten.beier@iwr.uni-heidelberg.de 
$^2$fred.hamprecht@iwr.uni-heidelberg.de
$^3$kappes@math.uni-heidelberg.de
}

\maketitle
\let\thefootnote\relax\footnote{This is an extended abstract. The full paper is available at the \href{http://www.cv-foundation.org/openaccess/CVPR2015.py}{Computer Vision Foundation webpage}. }
\vspace{-0.2in}

% Extended abstract begins here.  In a one-page document, there is
% little need for section headers, but you may use \section etc if you
% wish.

\noindent
Correlation clustering, or multicut partitioning,
is widely used in image segmentation for
partitioning an undirected graph or image with positive and negative edge weights 
such that the sum of cut edge weights is minimized.
%
Due to its NP-hardness, exact solvers do not scale and approximative solvers often give unsatisfactory results.
We investigate scalable methods for correlation clustering.
To this end we define fusion moves for the correlation clustering problem.

Our algorithm iteratively fuses the current and a proposed partitioning which 
monotonously improves
the partitioning and maintains a valid partitioning at all times.
Furthermore, it scales to larger datasets, gives near optimal solutions, and at the same time shows
a good anytime performance.

Correlation clustering~\cite{Bansal-2002}, also known as the multicut problem~\cite{chopra_1993_mp} 
is a basic primitive in computer vision~\cite{andres_2011_iccv,kroeger_2012_eccv,yarkony_2012_eccv,alush_2013_simbad} and data mining~\cite{Arasu-2009,Sadikov-2010,Chen-2012,Chierichetti-2014}.
%See Sec.~\ref{sec:problem_formulation} for its formal definition of clustering the nodes of a graph.

@@

\vspace{0.1cm}
\noindent \textbf{Contribution.}
In this work we present novel approaches that are designed for large scale correlation clustering problems.
First, we define a novel energy based agglomerative clustering algorithm that monotonically increases the energy.
With this at hand we show how to improve the anytime performance of Cut, Clue \& Cut~\cite{beier_2014_cvpr}.
%
Second, we improve the anytime performance of polyhedral multicut methods~\cite{kappes_2013_arxiv} by more efficient separation procedures.
%
Third, we introduce cluster-fusion moves, which extend the original fusion moves~\cite{Lempitsky-2010} 
used in supervised segmentation to the unsupervised case and give a polyhedral interpretation of this algorithm.
Finally, we propose two versatile proposal generators, and evaluate the proposed methods on existing and new benchmark problems.
Experiments show that we can improve the computation time by one to two magnitudes without worsening the segmentation 
quality significantly.

\bibliography{egbib}

\end{document}
